---
title: "Markdown refresh"
author: "David W."
date: "2024-06-09"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```   
# Section 1: C++ Introduction
## What is C++

C++ is a general purpose programming language that is free-form and compiled. It is regarded as an intermediate-level language, as it comprises both high-level and low-level language features. It provides imperative, object-oriented and generic programming features.

The overall program has a structure, but it is also important to understand the purpose of part of that structure.

## 101 step by step

```{r, eval = FALSE}
#include<iostream>

using namespace std;

int main()  {
    // YOUR CODE GOES HERE
    // Please take input and print output to standard input/output (stdin/stdout)
    // E.g. 'cin' for input & 'cout' for output
    cout << "Hello, InterviewBit!" << endl;
    return 0;
}
```
## preprocessor 
```{r, eval = FALSE}
#include<iostream>
```
- The hash sign (#) signifies the start of a preprocessor command. 
- The #include command is a specific preprocessor command that effectively copies and pastes the entire text of the file specified between the angle brackets into the source code. In this case, the file is “iostream” which is a standard file that should come with the C++ compiler. This file name is short for “input-output streams”; in short, it contains code for displaying and getting the text from the user.

- The include statement allows a programmer to “include” this functionality in the program without having to literally cut and paste it into the source code every time. 
- The iostream file is part of the C++ standard library, which provides a set of useful and commonly used functionality provided with the compiler. The “include” mechanism, however, can be used both for standard code provided by the compiler and for reusable files created by the programmer.

## namespace
```{r, eval = FALSE}
using namespace std;
```
C++ supports the concept of namespaces. A namespace 
- is essentially a prefix that is applied to all the names in a certain set. 
- like toolboxes with different useful tools. 

The using command tells the compiler to allow all the names in the “std” namespace to be usable without their prefix. 

The iostream file defines three names used in this program 
- cout, cin, and endl 

Those three are all defined in the std namespace. “std” is short for “standard” since these names are defined in the standard C++ library that comes with the compiler.

**Without using the std namespace, the names would have to include the prefix and be written as std::cout, std::cin, and std::endl.**

## main 
```{r, eval = FALSE}
int main()  {
```

The starting point of all C++ programs is the main function. 
This function is called by the operating system when your program is executed by the computer.

## the three command in std namespace     
```{r, eval = FALSE}
cout << "Hello, InterviewBit!" << endl;
```

**cin & cout**
- cout is character output
- cin is character input

In a typical C++ program, most function calls are of the form object.function_name(argument1, argument2).

Symbols such as **<<** can also behave like functions, as illustrated by the use of cout above. This capability is called operator overloading.

**{ }**
A block of code is defined with the { } tokens.

**semicolons**
Statements in C++ **must be terminated with a semicolon.**

## return
The return keyword tells the program to return a value to the **function int main** that called this function,  
and then to continue execution in the int main function from the point at which this function was called.

The **type of the value returned by a function** must match **the type specified in the declaration** of the function.

## Common data types in cpplus
built-in data types are described as follows:

- Int ("%d"): 32 Bit integer
- Long ("%ld"): 64 bit integer,  from -9,223,372,036,854,775,808 to 9,223,372,036,854,775,807
- Char ("%c"): Character type,  can range only from -128 to 127
- Boolean (either true or false)
- Float ("%f"): 32 bit real value
- Double ("%lf"): 64 bit real value

User Defined data types are 
- structures (struct) 
- classes (class). This will be covered later

## set precision for float and double
To print float and double up to specific number of decimal places use:

cout<<std::fixed << std::setprecision(number of decimal places)
- setprecision()  is available in the **iomanip** library.

## Type Modifiers
The above types can be modified using the following type modifiers: 
- signed 
- unsigned 
- short 
- long

## Typedefs
Creating new names (i.e. aliases) for existing types. 
Following is the simple syntax to define a new type using typedef:
```{r, eval = FALSE}
typedef int item
item number = 0 // number is a integer variable
```
As we have use typedef and created new name for int as item, item is equivalent to int now

## Variables
a variable is a container (storage area) to hold data.

int val = 10;
val = 15;

## Exercise 1.2
Try the following example in the editorial below.
You have to input 5 space-separated values: int, long, char, float and double respectively.

Print each value on a new line in the same order it is received as input. Note that the floating point value should be correct up to 3 decimal places and the double to 9 decimal places.

Example Input:

5 1234567891231 z 24.23 1214523.028352 
```{r, eval = FALSE}
#include<iostream>
#include <iomanip> 
using namespace std;

int main()  {
    int int_1; long long_1; char char_1; float float_1; double double_1; 
    cin >> int_1;
    cin >> long_1;   
    cin >> char_1;  
    cin >> float_1;  
    cin >> double_1;  
    
    cout << int_1 << endl << long_1 << endl << char_1 << endl << fixed << setprecision(3) << float_1 << endl << fixed << setprecision(9) << double_1 << endl;
        
    return 0;
}
```

## Type Conversion
C++ allows us to convert data of one type to that of another. This is known as type conversion.

There are two types of type conversion in C++.

Implicit Conversion
Explicit Conversion (also known as Type Casting)

## Implicit Type Conversion
The type conversion that is done automatically done by the compiler is known as implicit type conversion. This type of conversion is also known as automatic conversion.
```{r, eval = FALSE}
int a = 10;
char b = 'A';
a = a + b; // y implicitly converted to int. ASCII value of 'A' is 65
cout<<a<<endl;
// value of a is 75(10 + 65)

int num_int;
double num_double = 10.79;  
// implicit conversion  
// assigning a double value to an int variable  
num_int = num_double;  
// Value of num_int will be 10
// Here, the double value is automatically converted to int by the compiler before it is assigned to the num_int variable.
```

## Explicit Conversion
When the user manually changes data from one type to another, this is known as explicit conversion. This type of conversion is also known as type casting.

There are two major ways in which we can use explicit conversion in C++. They are:

- C-style type casting (also known as cast notation)
- type conversion operators

## C-style Type Casting
defining the required type **in front of the expression in parenthesis, i.e. (data_type)expression**. This can be also considered as forceful casting.

(data_type)expression;

```{r, eval = FALSE}
int a = 10;
char b = 'A';
a = a + (int)b;
cout<<a<<endl;
// value of a is 75
```

## Type conversion operators
C++ also has four operators for type conversion:
- static_cast:  static_cast<int>(variable)
- dynamic_cast
- const_cast
- reinterpret_cast

```{r, eval = FALSE}
float f = 4.5;   
// using cast operator 
int b = static_cast<int>(f); 
cout << b; 
// value of b is 4
To learn more about typecasting click http://www.cplusplus.com/doc/tutorial/typecasting/
```
## Exercise 1.3
You are given a character called ch, print the ASCII value of the character.
```{r, eval = FALSE}
#include<iostream>

using namespace std;

int main()  {
    char ch;
    cin>>ch;
    //Your code goes here
    int a;
    a =  int(ch);
    cout << a << endl;
    return 0;
}
```

## Math Library
The C++ math library is actually C’s math library. It is easy to use and is accessed by including cmath.
```{r, eval = FALSE}
#include <cmath>
```
Note: Trigonometric functions in cmath use RADIANS.

## Exercise 1.4
You are given two float variables A and B, perform the operations defined in comments in the editor below.

```{r, eval = FALSE}
#include<iostream>
#include<cmath>
using namespace std;
int main()  {
    float A = 12.56, B = 5.12;
    // Print the sum of cube of both A and B, and store it in float variable named "cube_val"
    float res_1 = pow(A,3) + pow(B,3);
    cout <<  res_1 << endl;
    
    // Print the square root of cube_val, and store it in float variable named "sq_val"
    float result_2 = sqrt(res_1);
    cout << result_2 << endl;
    
    // Print the sin of sq_val
    cout << sin(result_2) << endl;
    
    return 0;
}
``` 

## File Handling

Files are used to store data in a storage device permanently. 
File handling provides a mechanism to store the output of a program in a file, and to perform various operations on it.

In C++, files are mainly dealt by using three classes **fstream, ifstream, ofstream** available in the headerfile **fstream**.

- ofstream: Stream class, to write on files
- ifstream: Stream class, to read from files
- fstream: Stream class, to both read and write from/to files.

## Opening a file

```{r, eval = FALSE}
void open(const char* file_name,ios::openmode mode);
```
The first argument of the open function **const char* file_name)** defines the name, and format of the file with the address of the file.
The second argument **ios::openmode mode** represents the mode in which the file has to be opened. 

- in: File open for reading: the internal stream buffer supports input operations.
- out:	File open for writing: the internal stream buffer supports output operations.
- binary:	Operations are performed in binary mode rather than text.
- ate:	The output position starts at the end of the file.
- app:	All output operations happen at the end of the file, appending to its existing contents.
- trunc:	Any contents that existed in the file before it is open are discarded.

## Default Open Modes:
ifstream       ios::in
ofstream       ios::out
fstream        ios::in | ios::out
Note: You can combine two or more of these values by | them together.

```{r, eval = FALSE}
ofstream outfile;
outfile.open("file.dat", ios::out | ios::trunc );
```

Example using ifstream & ofstream classes.
```{r, eval = FALSE}
#include <iostream> 
#include <fstream> 
  
using namespace std; 
  
// Driver Code 
int main() 
{ 
    ofstream fout; // Creation of ofstream class object 
    string line; 
  
    // by default ios::out mode, automatically deletes 
    // the content of file. To append the content, open in ios:app 
    // fout.open("sample.txt", ios::app) 
    fout.open("sample.txt"); 
  
    // Execute a loop If file successfully opened 
    while (fout) { 
        getline(cin, line);  // Read a Line from standard input 
  
        if (line == "-1") // Enter -1 to exit 
            break; 
      
        fout << line << endl; // Write line in file 
    } 
    fout.close(); // Close the File 
  
    
    ifstream fin; // Creation of ifstream class object to read the file 
    fin.open("sample.txt");  // by default open mode = ios::in mode 
  
    // Execute a loop until EOF (End of File) 
    while (fin) { 
        getline(fin, line); // Read a Line from File
        cout << line << endl;  // Print line in Console 
    } 
  
    // Close the file 
    fin.close(); 
    return 0;
}
```

## Random Numbers
**rand()**
- rand() function is used to generate random numbers. 
```{r, eval = FALSE}
int rand(void): 
cout<<rand()<<endl; // Outputs any random number
```
returns a pseudo-random number in the range of 0 to RAND_MAX.
RAND_MAX: is a constant whose default value may vary between implementations but it is granted to be at least 32767.

**srand()**
sets the starting point for producing a series of pseudo-random integers. 
- If srand() is not called, the rand() seed is set as if srand(1) were called at program start. 
```{r, eval = FALSE}
void srand( unsigned seed ): 
```
Note: The pseudo-random number generator should only 
- be seeded once, 
- before any calls to rand(), 
- and the start of the program. 

It should not be 
- repeatedly seeded, or 
- reseeded every time you wish to generate a new batch of pseudo-random numbers.

Standard practice is to use the result of a call to **srand(time(0))** as the seed. 
```{r, eval = FALSE}
int main(){ 
    // This program will create different sequence of  
    // random numbers on every program run  
  
    // Use current time as seed for random generator 
    srand(time(0)); 
  
    for(int i = 0; i<4; i++) 
        cout<<rand()<<" "; 
    return 0; 
} 
```

# Section 2: Flow Control
## Comparision Operation & If-Else
- comparison operators such as ** ==, !=, >, <, ** etc can be used in C++

- These operators (cause the immediate statement in which they are contained to) return a Boolean value of either true or false.

Comparision Operator for all primitive data type (int, char, float, bool, etc.):
**== and != **

Comparision Operator for numeric data types only (int, float, double etc.):
**>, <, >=, <= **

## Conditional statements 
2 general types
- if, else / else if, else
- "switch…case" construct 

## Example 2.1:
Given an integer num denoting percentage of a student, calculate the grade according to the below rules:

If num >= 90, grade A.
If num >= 80, grade B.
If num >= 70, grade C.
If num >= 60, grade D.
If num >= 50, grade E.
Else grade will be F.
Print a string consisting of single character denoting the grade.
```{r, eval = FALSE}
#include<iostream>

using namespace std;

int main()  {
    int num;
    cin>>num;
    // Your code goes here
    char result;
    if (num >= 90){
        result = 'A';
    //console.log("A")
    }
    else if (num >= 80){
        result = 'B';
    }
    else if (num >= 70){
        result = 'C';
    }
    else if (num >= 60){
        result = 'D';
    }
    else if (num >= 50){
        result = 'E';
    }
    else {
        result = 'F';
    }
    cout << result << endl;

    return 0;
}
```
## Switch statement
```{r, eval = FALSE}
switch (expression) {
    case constant1:
        // code to be executed if expression is equal to constant1;
        break;

    case constant2:
        // code to be executed if expression is equal to constant2;
        break;

    default:
        // code to be executed if
        // expression doesn't match any constant
}
```
If there is a match, the corresponding code after the matching label is executed until the break statement is encountered.
```{r, eval = FALSE}
int x = 2; 
switch (x) 
{ 
    case 1: 
        cout << "Choice is 1"; 
        break; 
    case 2: 
        cout << "Choice is 2"; 
        break; 
    case 3: 
        cout << "Choice is 3"; 
        break; 
    default: 
        cout << "Choice other than 1, 2 and 3"; 
        break; 
}
```

## Exercise 2.2

```{r, eval = FALSE}
#include<iostream>

using namespace std;

int main()  {
    int weekday;
    cin>>weekday;
    // YOUR CODE GOES HERE
    string str_wkd;
    int numeric_input = int(weekday);
    switch (numeric_input){
        case 1: 
            str_wkd = "Monday";
            break;
        case 2:
            str_wkd = "Tuesday";
            break;
        case 3:
            str_wkd = "Wednesday";
            break;
        case 4:
            str_wkd = "Thursday";
            break;
        case 5: 
            str_wkd = "Friday";
            break;
        case 6:
            str_wkd = "Saturday";
            break;
        case 7:
            str_wkd = "Sunday";  
            break;
    }
    cout << str_wkd << endl;
    return 0;
} 
```

## Conditional or Ternary Expression
```{r, eval = FALSE}
( condition ) ? expressionIfTrue : expressionIfFalse;
```
First the condition is evaluated, and the side effects of this evaluation carry out their impact on the local environment.

If the result is true then only the **expressionIfTrue** is evaluated (causing side effects) and this second result is the value of the whole conditional expression, and the expressionIfFalse is not evaluated (and hence cause no side effects).

If the condition evaluates to false, then the situation is converse

A common use of the conditional expression:
- assign the value x or y to a, depending on an easily decidable condition, say x > y.
```{r, eval = FALSE}
int x = 7;
int y = 5;
int a = ( x > y ) ? x : y;
cout << a << endl;
```

## Exercise 2.3
Given two integer x and y, using Conditional or Ternary expression, print “Robin” if the value of x is less than or equal to y else print “Rahul” (without quotes).

```{r, eval = FALSE}
#include<iostream>

using namespace std;

int main()  {
    int x, y;
    cin>>x>>y;
    // YOUR CODE GOES HERE
    string ans = ( x <= y )? "Robin" : "Rahul";

    cout << ans << endl;

    return 0;
}
```
## Loops
3 Types of loop:
1. for 
2. while
3. do...while

## for loop
for (initialization; condition; update) {
    // body of-loop 
}
for(int i = 1; i <= 5; i++){
   cout<<i<<" ";
}
1. initialization - initializes variables and is executed only once
2. condition - if true, the body of for loop is executed, if false, the for loop is terminated
3. update - updates the value of initialized variables and again checks the condition

## while loop
while (condition) {
    statememt(s);
}
int i = 1;

while(i <= 5){
    cout<<i<<" ";
    i++;
}

## do…while loop
- do…while loop checks the conditional statement after the first run, then continuing onto another iteration.
- A do-while loop is used where your loop should execute at least one time.
Ex:  take an integer input from the user until the user has entered a positive number. 

do {
  //body
} while (condition);

int i = 1;
do {
    cout << i << " ";
    i++;
} while(i <= 5);   // the contition is being checked after the first run

## Exercise 2.4

```{r, eval = FALSE}
#include<iostream>

using namespace std;

int main()  {
    int N;
    cin>>N;
    // YOUR CODE GOES HERE
    int i = 0;
    while( i <= N) {
        if(i % 2 != 0) {
            cout << i << endl;
        }
        i++;
    }
    return 0;
}
```

## Jump Statements
Jump statements are used to **interrupt the normal flow** of program.

Types of Jump Statements:
1. break
2. continue
3. goto

## Break Statement
- used inside loop, or switch statement. 
- compiler will abort the loop and continue to execute statements followed by loop.
int a = 1;  

while(a <= 10)  
{  
    if(a==5)  
        break;  
    a++;  
}  
cout << "Value of a is " <<a<<endl;  
// Value of a is 5

## Continue Statement
- inside loop
- skip the followling statements in the current loop, and resume the loop

int a = 0;  
while(a < 10)  
{  
    a++;  
    if(a == 5)  
        continue;  

    cout << a << " ";  
}  
// prints 1 2 3 4 6 7 8 9 10 

## Goto Statement
- jumps from one point **to another point within a function**.
goto label_1;  (a)
... .. ...  
... .. ...  
... .. ...  
label_1:      (b)
statement;  
... .. ...
- When **goto label_1;** (a) is encountered, the control of program jumps to **label_1:** (b) and executes the code below it.

num = 10      
if (num % 2 == 0)   
    // jump to even   
    goto even;    
else  
    // jump to odd   
    goto odd;    
even:   
    cout << num << " is even";   
    // return if even   
    return;    
odd:   
    cout << num << " is odd"; 

## Reason to Avoid goto Statement
- can write any C++ program without the use of goto statement
- makes the logic of the program complex and tangled
- a harmful construct and a bad programming practice

## Exercise 2.5

You are given an integer N, print all the odd values, for all i, where 0 <= i < N. Print each values on a seperate line.
Note: Use continue statement

```{r, eval = FALSE}
#include<iostream>

using namespace std;

int main()  {
    int N;
    cin>>N;
    // YOUR CODE GOES HERE
    for(int i = 0; i <= N; i++) {
        if (i % 2 == 0) {
            continue;
        }
        cout << i << endl;
    }
   
    return 0;
}
```